Gusto: a toolkit for compatible finite element dynamical cores

### 2024-03-07 source/paper.tex
better start for this paragraph describing the software structure

\JScomment{Text here depends on what the schematic looks like - this
  paragraph should refer to the schematic and explain all aspects}
Prognostic equation sets are each defined in a python class, making
use of the common finite element forms for the transport and
difffusion terms. The timeloop class contains the definition of the
timestepping procedure and the \texttt{run} method that loops over the
timesteps. It also sets up diagnostic fields and calls the IO as
required. The simplest timeloop class will just apply one time
discretisation method to the equation set but it is common in weather
and climate prediction models to split the different types of terms
and apply a time discretisation separately to each type of term. For
example, the transport and forcing terms might be solved for
separately, such as in **, which enables explicit transport schemes to
be applied to each field. ** something about SIQN structure. Terms
related to the physics are solved for in the appropriate part of the
iteration loop, with physics processes that occur on `fast' timescales
evaluated more frequently and `slower' physics processes evaluated
just once at the end of the timestep. The choice of where to evaluate
the physics processes in the iteration is also based on the
computational cost and the best way to do this is still an open
research question (**ref).

@@

\section{Form Manipulation Labelling}
\label{sec: FML}
The discussion of the time discretisation so far has been general i.e.
not specific to any particular equation set. This is also true of the
implementation of the timestepping schemes. This enables flexibility
and reproducibility (**elaborate). This functionality is enabled
through form manipulation labelling (FML) which is used to associate a
label to each term in the residual form defined in the equation
class. The timestepping methods are then able to reason about which
manipulations to apply to which forms by applying filters. A simple
example would be a splitting method that evaluates transport terms and
diffusion terms separately. The full residual is passed to the
transport scheme which then applies a filter to drop all terms that
are not labelled as transport; similarly for diffusion. We illustrate
this procedure in figure \ref{fig:advdiff} where we show the
definition of the residual for the advection diffusion equation for a
scalar variable $q$ with a prescribed transport velocity $\MM{c}$ and
diffusivity $\kappa$

>>>>

\begin{figure}
  \centering{\includegraphics[width=0.8\textwidth]{gusto_schematic.png}}
\end{figure}

\JScomment{Text here depends on what the schematic looks like - this
  paragraph should refer to the schematic and explain all aspects} The
Gusto model comprises a domain and a prognostic equation set. The
domain holds the mesh and the finite element function spaces along
with any other information about the geometry such as the rotation
vector. The Prognostic equation sets are each defined in a python
class, making use of the common finite element forms for the transport
and difffusion terms. The timeloop class contains the definition of
the timestepping procedure and the \texttt{run} method that loops over
the timesteps. It also sets up diagnostic fields and calls the IO as
required. The simplest timeloop class will just apply one time
discretisation method to the equation set but it is common in weather
and climate prediction models to split the different types of terms
and apply a time discretisation separately to each type of term. For
example, the transport and forcing terms might be solved for
separately, such as in **, which enables explicit transport schemes to
be applied to each field. ** something about SIQN structure. Terms
related to the physics are solved for in the appropriate part of the
iteration loop, with physics processes that occur on `fast' timescales
evaluated more frequently and `slower' physics processes evaluated
just once at the end of the timestep. The choice of where to evaluate
the physics processes in the iteration is also based on the
computational cost and the best way to do this is still an open
research question (**ref).

@@

\section{Form Manipulation Language}
\label{sec: FML}
The discussion of the time discretisation so far has been general i.e.
not specific to any particular equation set. This is also true of the
implementation of the timestepping schemes. This enables flexibility
and reproducibility (**elaborate). This functionality is enabled
through the form manipulation language (FML) which is used to
associate a label to each term in the residual form defined in the
equation class. The timestepping methods are then able to reason about
which manipulations to apply to which forms by applying filters. A
simple example would be a splitting method that evaluates transport
terms and diffusion terms separately. The full residual is passed to
the transport scheme which then applies a filter to drop all terms
that are not labelled as transport; similarly for diffusion. We
illustrate this procedure in figure \ref{fig:advdiff} where we show
the definition of the residual for the advection diffusion equation
for a scalar variable $q$ with a prescribed transport velocity
$\MM{c}$ and diffusivity $\kappa$
